---
title: "Reproducible Research: Peer  Assessment 1"
subtitle: "Analysis of movement of data collected of one individual by activity monitoring device"
author: CSoaresF   
date: Jan 09, 2016
output: html_document:   
        keep_md: true   
--- 
## Analysis of data collected of movement of one individual by activity monitoring device  
#### CsoaresF   
=========================================  
## Introduction
Analysis of data about personal movement using activity monitoring devices.   
The device collects dat at 5 minute intervals through out the day.   
The data consists of two months of data an anonymous individual collected during the months of October and November 2012 and include the number of steps taken in 5 minute interval each day.      
   
#### Dataset: Activity monitoring data [52k]   
*  The dataset is stored in a comma-separated-value (CSV) file.   
*  There are a total of 17,568 observations in this dataset.
*  url for download: https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip   
    
*  Variables of dataset:
    +  __steps__: Number of steps in a 5 minute interval   
          (missing values are coded as NA)   
    +  __date__: The date on wich the measurement was taken   
          format: YYYY-MM-DD   
    +  __interval__: Identifier for the 5 minute interval in which measurement was taken   


## 1- Loading and preprocessing the data   
```{r echo = TRUE}
# libraries
library(lattice)
library(ggplot2)
library(plyr)
library(dplyr)
#################

# work directory
setwd("C:/COURSERA5")
# if file is not exist in work directory, download and unzip
if(!file.exists("activity.csv")) {
    file <- "activity.zip"
    web <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
    download.file(web, file)
    unzip(file)
}

# load spreadsheet to work space
activity <- read.csv("activity.csv")

# conversion to date
activity$date <- as.Date(as.character(activity$date), "%Y-%m-%d")

#  view file structure
str(activity)
```
## 2- What is mean total number of steps taken per day?   
```{r echo = TRUE}
# total steps by day (using dplyr)
tot.steps.day <- activity %>%                    # read activity,
  group_by(date) %>%                             # group by date,
  summarise(steps = sum(steps, na.rm=FALSE)) %>% # summarise steps (with NAs),
  arrange(date)                                  # and order by date.

# view file struture
str(tot.steps.day)
# number of days in dataset
length(tot.steps.day$date)

# total steps by day (with NAs)
summary(tot.steps.day)
```
*  The mean is 10,770 steps by day.   
*  The median is 10,760 steps by day.   
*  Ignoring missing values in the data, mean and medin are very close.   
   
```{r echo = TRUE}
# histogram (with NAs)
qplot(tot.steps.day$steps, 
      geom = "histogram",
      binwidth = 700) +
  labs(title="Histogram steps by day") +
  labs(x="Quantity steps by day") +
  labs(y="Count days")

#
# Quantity of steps per day, one person, for 61 days (Oct/1/2012 - Nov/30/2012).
print(tot.steps.day$steps)
```

## 3- What is the average daily activity pattern?   
```{r echo = TRUE}
# 
# time series plot (type = "l")
# interval 5min: x-axis
# mean of steps taken: y-axis
mean.steps <- aggregate(steps ~ interval, data = activity, FUN = mean)
plot(mean.steps, type = "l")
```
#
## 4- Imputing missing values   
```{r echo = TRUE}
# 
summary(activity)
summary(tot.steps.day)
#######################################
# total steps by day without NAs
tot.steps.day2 <- activity %>%                  # read activity,
  group_by(date) %>%                            # group by date,
  summarise(steps = sum(steps, na.rm=TRUE)) %>% # summarise steps (without NAs),
  arrange(date)                                 # and order by date.

# 
summary(tot.steps.day2)
############################ 4444444444
steps.by.day <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)
############################## 44444444
# missing values in activity dataset
missing <- !complete.cases(activity)
sum(missing == TRUE)
# 
activity2 <- activity[missing == TRUE, ]
steps.by.interval <- aggregate(steps ~ interval, data = activity, mean, na.rm = TRUE)
activity2[, 1] <- steps.by.interval$step
# activity3: dataset with missing data filled in 
activity3 <- rbind(activity[complete.cases(activity), ], activity2)
summary(activity2)
summary(activity3)
#
# total steps by day with missing data filled in
steps.by.day3 <- aggregate(steps ~ date, data = activity3, sum)
hist(steps.by.day3$steps)
summary(steps.by.day3)

```
#
## 5- Are there differences in activity patterns between weekdays and weekends?   
   
Weekends: Saturday (sáb), Sunday (dom).    
Weekdays: Mon (seg), Tues (ter), Wed (qua), Thurs (qui), Fri (sex).   
```{r echo = TRUE}
# 
activity3$day.week <- weekdays(activity3$date, abbreviate = TRUE)
head(activity3)
tail(activity3)
#
activity3$day.week[activity3$day.week == "sáb" | activity3$day.week == "dom"] <- "weekend"
activity3$day.week[activity3$day.week != "weekend"] <- "weekday"
activity3$day.week <- as.factor(activity3$day.week)
head(activity3)
str(activity3)
#
# painel plot weekday x weekend
# interval 5m x-axis
# mean steps y-axis
activity4 <- ddply(activity3, .(interval, day.week), summarise, mean.steps = mean(steps))
summary(activity4)
#
xyplot(mean.steps ~ interval | day.week, data = activity4, type = "l", lwd = 2, layout = c(1, 2), ylab = "Number of steps")
```
